---
title: "p8105_hw2_iah2124redo"
author: "Iris Hart"
date: "2024-10-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

### Problem 0: Loading Packages 
```{r load_libraries}
library(tidyverse)
library(readxl)
```

### Problem 1
```{r}
trans_ent = 
  read_csv(
    "NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) |> 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

```{r}
dim(trans_ent)
```
This data set contains the variables: line, station name, station latitude, station longitude, routes, entry, exit only, vending, entrance type, and ada compliance. Cleaned the data by reading the file and using janitor to keep the variables needed, and change the names accordingly. The dimensions of this data set are 1868 x 20. This data is not considered "tidy". This is because route number should be a variable, as should route. That is, to obtain a tidy data set we would need to convert `route` variables from wide to long format. 

How many distinct stations are there?
```{r}
trans_ent |> 
  select(station_name, line) |> 
  distinct()
```

There are 465 distinct stations. 

How many stations are ADA compliant?
```{r}
trans_ent |> 
  filter(ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```

84 stations are ADA complicit. 

What proportion of station entrances / exits without vending allow entrance?
```{r}
trans_ent |> 
  filter(vending == "NO") |> 
  pull(entry) |> 
  mean()
```

37.71% of station enterances / exits without vending allow entrance.

```{r}
trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A") |> 
  select(station_name, line) |> 
  distinct()

trans_ent |> 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") |> 
  filter(route == "A", ada == TRUE) |> 
  select(station_name, line) |> 
  distinct()
```
60 stations service the A train, 17 of them being ADA compliant. 

### Problem 2
Reading the Mr. Trash Wheel data from the 2024 Trash Wheel Collection Data
```{r}
mtw = read_excel("202409 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:M"))
```
Editing variable names to be unique
```{r}
names(mtw) <- make.names(names(mtw), unique = TRUE)
```
Removing the last two lines of the sheet, no data contained there
```{r}
mtw = mtw[-c(652, 653), ]
```
Viewing the cleaned Mr. Trash Wheel data
```{r}
head(mtw)
```

Rounding the number of sports balls to the nearest integer 
```{r}
mtw <- mtw %>%
  mutate(Sports.Balls = as.integer(round(as.numeric(Sports.Balls))))
```
Check for rounded value in sports balls column
```{r}
head(mtw)
```

Reading the Professor Trash Wheel data from the 2024 Trash Wheel Collection Data
```{r}
ptw = read_excel("202409 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = cell_cols("A:L"))
```

Editing variable names to be unique
```{r}
names(ptw) <- make.names(names(ptw), unique = TRUE)
```

Removing the last three lines of the sheet, no data contained there
```{r}
ptw = ptw[-c(119, 120, 121), ]
```

Viewing the cleaned Professor Trash Wheel data
```{r}
head(ptw)
```

Reading the Gwynnda Trash Wheel data from the 2024 Trash Wheel Collection Data
```{r}
gtw = read_excel("202409 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = cell_cols("A:K"))
```

Editing variable names to be unique
```{r}
names(gtw) <- make.names(names(gtw), unique = TRUE)
```

Removing the last line of the sheet, no data contained there
```{r}
gtw = gtw[-264, ]
```

Viewing the cleaned Gwynnda Trash Wheel data
```{r}
head(gtw)
```

Combining data sets 
```{r}
mtw <- mtw %>% mutate(Trash_Wheel = "Mr. Trash Wheel")
ptw <- ptw %>% mutate(Trash_Wheel = "Professor Trash Wheel")
gtw <- gtw %>% mutate(Trash_Wheel = "Gwynnda Trash Wheel")
```

Viewing structure of data set variables
```{r}
str(mtw)
str(ptw)
str(gtw)
```
Converting date variable into character 
```{r}
mtw <- mtw %>% mutate(Year = as.character(Year))
ptw <- ptw %>% mutate(Year = as.character(Year))
gtw <- gtw %>% mutate(Year = as.character(Year))
```
Combining the data sets
```{r}
combined_data <- bind_rows(mtw, ptw, gtw)
```
Making sure that trash wheel is the first column in the data set
```{r}
combined_data <- combined_data %>%
  select(Trash_Wheel, everything()) %>%  # Ensure Trash_Wheel is the first column
  distinct()
```
Printing the combined data
```{r}
print(combined_data)
```
Identifying how many total rows are in the dataset
```{r}
nrow(combined_data)
```
Identifying the total weight of trash collected by Professor Trash Wheel
```{r}
total_weight_professor_trash_wheel <- sum(ptw %>% pull(Weight..tons.), na.rm = TRUE)
```
Printing the total weight of trash collected by Professor Trash Wheel
```{r}
print(total_weight_professor_trash_wheel)
```

Identifying the total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
total_cigarette_butts_gwynnda <- sum(gtw %>% filter(Month == "June", Year == 2022) %>% pull(Cigarette.Butts), na.rm = TRUE)
```
Printing the total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
print(total_cigarette_butts_gwynnda)
```
Key variables
```{r}
glimpse(combined_data)
```

In this combined data set containing data from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda Trash Wheel sets, there are 1032 total observations. Key variables in this data set are which trash wheel the data comes from, month,  year, the weight (in tons), date of trash collection, and cigarette butts collected. The total weight of trash collected by Professor Trash Wheel is 246.72tons. The total number of cigarette butts collected by Gwynnda in June of 2022 was 18120. This combined data set shows key information on how much trash is collected by each wheel, type of trash, and when it was collected. 


### Problem 3 
Importing and cleaning the data
```{r}
bakers_df = read_csv("gbb_datasets/bakers.csv") |>
  janitor::clean_names() 

bakes_df = read_csv("gbb_datasets/bakes.csv",
                    na = c("NA", "", ".", "N/A", "UNKNOWN", "Unknown")) |>
  janitor::clean_names() 

results = read_csv("gbb_datasets/results.csv", skip = 2) |>
   janitor::clean_names() 
```
Change variable to be in unison, baker for the baker name across all datasets
```{r}
bakers_df = bakers_df |>
  separate(baker_name, into = c("baker", "baker_last_name"), sep = " ")
```
Check for completeness 
```{r}
na_bakers_df = bakers_df |>
  anti_join(bakes_df, by = c("baker", "series"))
```

```{r}
print(na_bakers_df)
```
25 reported missing bakers who do not have the information across datasets

Checking for those who do not have the same information recorded in bakers and bakes data sets
```{r}
anti_join(bakers_df, bakes_df)
```

Checking for those who do not have the same information in bakers and results data sets
```{r}
anti_join(bakers_df, results)
```
Identified Jo Wheatley 

Change Jo's name to be consistent across data sets
```{r}
results = results |>
  mutate(baker = case_match(baker, "Joanne" ~ "Jo"))
```